---
title: "Track COVID-19 Literature"
author: "Halie Rando"
date: "4/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set
source("CDC_query_functions.R")
```

## Pull Up-To-Date List of Papers

Run the code below to query the CDC website for the most up-to-date list of papers related to COVID-19.

```{R updatedb}
# Go to CDC website and download the most recent list of papers
COVID19_data <- getdata()
print(paste("Database last updated at", Sys.time(), "and contains", 
            nrow(COVID19_data), "papers"))
```

## Identify Papers for Each Therapeutic

Below, for each therapeutic, call the `findpapers` function.
This will search through the most recent CDC list for all papers containing your keywords of interest.
You must specify a vector of keywords (in R, a vector is what we'd conversationally call a list).
The therapeutic itself should be the first keyword.
The format will look like this:
`findpapers(c("Remdesivir", "GS-5734", "GS-441524", "adenosine"))`
We are assuming each keyword is one word and < 100 characters. 
Please don't use any unnecessary capitalization.
Each call of findpapers() will generate a separate csv output file.

```{r grabkwds}
findpapers(c("Remdesivir", "GS-5734", "GS-441524", "adenosine"))
```

Now navigate to the [`bytopic`](bytopic) folder to explore each therapeutic individually.